\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-04-13 write_up/mixfishsim_ecolmod.tex
Discussion

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time with varying knowledge of species distributions
using species-unselective fishing gear. \added[id=PD]{In doing so} fisheries
\deleted[id=PD]{that}catch an assemblage of species
\added[id=PD]{and}\deleted{, known as mixed fisheries} \replaced[id=JJ]{may}{,
	when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} overquota catch \added[id=JJ]{when managed by
	single species quotas and fishers exhaust or more quota, may} lead to
overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{This discarding of fish in
	excess of quota hampers the ability to limit fishing mortality to
	within sustainable limits \citep{Alverson1994, Crowder1998,
		Rijnsdorp2007}} \deleted[id=PD]{; reducing discarding is
	crucial}and the ability to manage for the biological and economic
sustainability of fisheries\deleted[id=JJ]{ and implementation of an ecosystem
	approach to fisheries}. As such, there is increasing interest in
technical solutions such as gear and spatial closures as ways of
\replaced[id=PD]{reducing unwanted catch}{avoiding discard\added[id=JJ]{ing of
		fish}}\citep{Kennelly2002, Catchpole2008, Bellido2011}.  \\

@@

The basic population level processes are simulated using a modified two-stage
Deriso-Schnute delay difference model \citep{Deriso1980, Schnute1985,
	Dichmont2003} occurring at a daily time-step. \added[id=PD]{A daily
	time-step was chosen as to discretise continuous population processes
	on a biologically relevant and computationally tractable timescale.}
\replaced[id=PD]{Under the population dynamics module}{Here,} population
biomass growth and depletion for pre-recruits and \deleted[id=PD]{fish}
recruited \added[id=PD]{fish} \deleted[id=PD]{to the fishery} are modelled
separately as a function of previous recruited biomass, intrinsic population
growth and recruitment. Biomass for each cell is incremented each day as
follows (the full parameter list is detailed in Table \ref{tab: pop_variable}):
\begin{equation*}
	\begin{split}
	B_{c,d+1} = &\\
	& (1 + \rho) B_{c,d} \cdot e^{-Z_{c,d}} - \rho \cdot e^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& (B_{c,d-1} \cdot e^{-Z_{c,d-1}} + Wt_{R-1} \cdot \alpha_{d-1} \cdot
	R_{\tilde{y}(c,y,d-1)})
	\hspace{0.4cm} + \\
	& Wt_{R} \cdot \alpha_{d} \cdot R_{\tilde{y}(c,y,d)} 
	\end{split}
\end{equation*}
where $\rho$ is Brody's coefficient, shown to be approximately equal to
$e^{-K}$ when $K$ is the growth rate from a von Bertalanffy logistic growth
model \citep{Schnute1985}. $Wt_{R-1}$ is the weight of fish prior to
recruitment, while $Wt_{R}$ is the recruited weight. $\alpha_{d}$ represents
the proportion of fish recruited during that day for the year, while
$R_{c,\tilde{y}}$ is the annual recruits in cell $c$ for year $y$. \\

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation*}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} * (1 - e^{-(F_{c,d} + M_{c,d})}) *
B_{c,d}
\end{equation*}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. Here, catch and fishing
mortality are the sum of those across all fleets and vessels, where $F_{fl, v,
	c, d, p} = E_{fl, v, c, d} \cdot Q_{fl, p} \cdot B_{c, d, p}$ with
$fl$, $v$ and $p$ the fleet, vessel and population respectively and $E$ and $Q$
fishing effort and catchability. \todo{\added[id=CM]{[link $F$ to effort and
		catchability - as I think we have F as an emergent property of
		the fleets rather than something we solve for (I could be wrong
		though!) - catch for a vessel is a product of catchability and
		biomass, i.e. C = qB, but this catch is summed to solve for F.
		So its both really]}}\\

@@

Recruitment is modelled through a function relating the mature biomass to
recruits at time of recruitment. In \emph{MixFishSim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
\citep{Beverton1957}: 
\begin{equation*}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha * B_{c,d})}{(\beta + B_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation*}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum, $B$ current SSB and
$\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation*}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} * e^{(\alpha - \beta * B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation*}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases. In this study, the
Beverton-Holt form of stock recruit relationship was used for all populations.

@@

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by \deleted[id=JJ]{relating} recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$),
\begin{equation}
Rev = \sum_{p=1}^{P} \replaced[id=PD]{L}{C}_{p} \cdot Pr_{p} 
\end{equation}
where $\replaced[id=PD]{L}{C}_{p}$ is \replaced[id=PD]{landings}{catch} of a
population $p$, and $Pr_{p}$ price of a population. Here, when fishing is
successful vessels remain in a similar location and continue to exploit the
local fishing grounds. When unsuccessful, they move some distance away from
the current fishing location.  The movement distance retains some degree of
stochasticity, which can be controlled separately, but is determined by the
relationship: 
\begin{equation*}
	StepL = e^{log(\beta_{1}) + log(\beta_{2}) -
		(log(\frac{\beta_{1}}{\beta_{3}}))} * Rev
\end{equation*}\todo{\added[id=JJ]{So
			step length increases with increasingly gross
			revenue?}\added[id=PD]{No, the opposite}}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
(x1,y1) to (x2, y2) is defined by:
\begin{equation*}
	\begin{split}
 (x2, y2) =  & x1 + StepL \cdot \cos (\frac{\pi \cdot Br}{180}), \\
             & y1 + StepL \cdot \sin (\frac{\pi \cdot Br}{180}) \\	
 with  \hspace{0.5cm}     & Br_{t-1} < 180, Br_{t} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t-1} > 180, Br_{t} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation*}
where $k$ the concentration parameter from the von \replaced[id=JJ]{M}{m}ises
distribution which we correlate with the revenue so that $k = (Rev + 1 /
RefRev) * max_{k}$, where $max_{k}$ is the maximum concentration value, $k$,
and $RefRev$ is parametrised as for $\beta_{3}$ in the step length function. A
realised example of the step length and turning angle relationships to revenue
can be seen at Figure \ref{S-fig:15}.

@@

The finer spatial grid for the real population (top left) and commercial data
(top middle) show similar patterns, though there are unsampled gaps in the
commercial data from a lack of fishing activity (particularly in the lower left
part of the sampling domain). The survey data at this spatial resolution shows
very sparse information about the spatial distributions of the populations. The
slightly aggregated data on a 5 x 5 grid shows similar patterns, and while
losing some of the spatial detail there remains good consistency between the
'real population' and the commercial data. Survey data starts to pick out some
of the similar patterns as the other data sources, but lacks coverage. The
spatial catch information on a 10 x 10 and 20 x 20 grid loses a significant
amount of information about the spatial resolutions for all data sources, and
some differences between the survey, commercial and 'real population' data
emerge. \\

@@

For the commercial data, the most effective closure scenario was based on 1 x 1
data at a monthly temporal resolution. This results in $\sim$ 10 \% reduction
in F for species 1. \textcolor{blue}{This was the only closure scenario to have
	positive effect according to Figure \ref{fig:4}, though looking at the
	trend in Figure \ref{fig:3} this looks more related to the continued
	increased in F trend, as other scenarios had an initial effect}.
Interestingly the monthly data scenario was more effective than weekly data,
which I'd posit is due to the increase amount of data available from the
commercial sampling across a month compared to a week.i Commercial data used at
an annual timestep was ineffective in bringing fishing mortality down for
species 1. \\

Given the scenarios above, it seems clear that spatial disaggregation is more
important than the temporal disaggregation of the commercial data, except when
its used at an annual timeframe, which is the scenario that gave the worst
results.

@@

\underline{The why:} \\

Dynamic management requires understanding of the interaction between fishers
and heterogeneously distributed resources. \\

Difficult because of a lack of knowledge of the scale interactions important at
and detailed spatiotemp knowledge. \\

\underline{The what:}  \\

We develop a simulation model to understand scale.\\

Detail of sim model important -- highly resolved, IBM. Make comparison to other
models, e.g. DISPLACE, POSEIDON...\\

@@

\underline{The what it shows:} \\

Our results show commercial data can be useful at right scale and resolution -
depends on scale of process: pop movement etc... Important to consider how
fishers interact / adapt to changes with the resoruce and mgmt. \\

@@

%%%% figures
\begin{figure}[!ht]
	\includegraphics[width =\linewidth]{../analysis/Data_Aggregation_space}
	\caption{Data aggregation at different spatial resolutions over a
		ten year period}
	\label{fig:1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../analysis/Data_Aggregation_time}
	\caption{Data aggregation at different temporal resolutions over a
		three year period}
	\label{fig:2}
\end{figure}

>>>>

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time with varying knowledge of species distributions
using species non-selective fishing gear. \added[id=PD]{In doing so} fisheries
\deleted[id=PD]{that}catch an assemblage of species
\added[id=PD]{and}\deleted{, known as mixed fisheries} \replaced[id=JJ]{may}{,
	when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} over-quota catch \added[id=JJ]{when managed by
	single species quotas and fishers exhaust or more quota, may} lead to
overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{This discarding of fish in
	excess of quota hampers the ability to limit fishing mortality to
	within sustainable limits \citep{Alverson1994, Crowder1998,
		Rijnsdorp2007}} \deleted[id=PD]{; reducing discarding is
	crucial}and the ability to manage for the biological and economic
sustainability of fisheries\deleted[id=JJ]{ and implementation of an ecosystem
	approach to fisheries}. As such, there is increasing interest in
technical solutions such as gear and spatial closures as ways of
\replaced[id=PD]{reducing unwanted catch}{avoiding discard\added[id=JJ]{ing of
		fish}}\citep{Kennelly2002, Catchpole2008, Bellido2011}.  \\

@@

The basic population level processes are simulated using a modified two-stage
Deriso-Schnute delay difference model \citep{Deriso1980, Schnute1985,
	Dichmont2003} occurring at a daily time-step. \added[id=PD]{A daily
	time-step was chosen as to discretise continuous population processes
	on a biologically relevant and computationally tractable timescale.}
\replaced[id=PD]{Under the population dynamics module}{Here,} population
biomass growth and depletion for pre-recruits and \deleted[id=PD]{fish}
recruited \added[id=PD]{fish} \deleted[id=PD]{to the fishery} are modelled
separately as a function of previous recruited biomass, intrinsic population
growth and recruitment. Biomass for each cell is incremented each day as
follows (the full parameter list is detailed in Table \ref{tab: pop_variable}):
\begin{equation}
	\begin{split}
	B_{c,d+1} = &\\
	& (1 + \rho) B_{c,d} \cdot e^{-Z_{c,d}} - \rho \cdot e^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& (B_{c,d-1} \cdot e^{-Z_{c,d-1}} + Wt_{R-1} \cdot \alpha_{d-1} \cdot
	R_{\tilde{y}(c,y,d-1)})
	\hspace{0.4cm} + \\
	& Wt_{R} \cdot \alpha_{d} \cdot R_{\tilde{y}(c,y,d)} 
	\end{split}
\end{equation}
where $\rho$ is Brody's coefficient, shown to be approximately equal to
$e^{-K}$ when $K$ is the growth rate from a von Bertalanffy logistic growth
model \citep{Schnute1985}. $Wt_{R-1}$ is the weight of fish prior to
recruitment, while $Wt_{R}$ is the recruited weight. $\alpha_{d}$ represents
the proportion of fish recruited during that day for the year, while
$R_{c,\tilde{y}}$ is the annual recruits in cell $c$ for year $y$. \\

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} * (1 - e^{-(F_{c,d} + M_{c,d})}) *
B_{c,d}
\end{equation}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. Here, catch and fishing
mortality are the sum of those across all fleets and vessels, where $F_{fl, v,
	c, d, p} = E_{fl, v, c, d} \cdot Q_{fl, p} \cdot B_{c, d, p}$ with
$fl$, $v$ and $p$ the fleet, vessel and population respectively and $E$ and $Q$
fishing effort and catchability. \todo{\added[id=CM]{[link $F$ to effort and
		catchability - as I think we have F as an emergent property of
		the fleets rather than something we solve for (I could be wrong
		though!) - catch for a vessel is a product of catchability and
		biomass, i.e. C = qB, but this catch is summed to solve for F.
		So its both really]}}\\

@@

Recruitment is modelled through a function relating the mature biomass to
recruits at time of recruitment. In \emph{MixFishSim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
\citep{Beverton1957}: 
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha * B_{c,d})}{(\beta + B_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum, $B$ current SSB and
$\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} * e^{(\alpha - \beta * B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases. In this study, the
Beverton-Holt form of stock recruit relationship was used for all populations.

@@

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by \deleted[id=JJ]{relating} recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$),
\begin{equation}
Rev = \sum_{p=1}^{P} \replaced[id=PD]{L}{C}_{p} \cdot Pr_{p} 
\end{equation}
where $\replaced[id=PD]{L}{C}_{p}$ is \replaced[id=PD]{landings}{catch} of a
population $p$, and $Pr_{p}$ price of a population. Here, when fishing is
successful vessels remain in a similar location and continue to exploit the
local fishing grounds. When unsuccessful, they move some distance away from
the current fishing location.  The movement distance retains some degree of
stochasticity, which can be controlled separately, but is determined by the
relationship: 
\begin{equation}
	StepL = e^{log(\beta_{1}) + log(\beta_{2}) -
		(log(\frac{\beta_{1}}{\beta_{3}}))} * Rev
\end{equation}\todo{\added[id=JJ]{So
			step length increases with increasingly gross
			revenue?}\added[id=PD]{No, the opposite}}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
(x1,y1) to (x2, y2) is defined by:
\begin{equation}
	\begin{split}
 (x2, y2) =  & x1 + StepL \cdot \cos (\frac{\pi \cdot Br}{180}), \\
             & y1 + StepL \cdot \sin (\frac{\pi \cdot Br}{180}) \\	
 with  \hspace{0.5cm}     & Br_{t-1} < 180, Br_{t} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t-1} > 180, Br_{t} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation}
where $k$ the concentration parameter from the von \replaced[id=JJ]{M}{m}ises
distribution which we correlate with the revenue so that $k = (Rev + 1 /
RefRev) * max_{k}$, where $max_{k}$ is the maximum concentration value, $k$,
and $RefRev$ is parametrised as for $\beta_{3}$ in the step length function. A
realised example of the step length and turning angle relationships to revenue
can be seen at Figure \ref{S-fig:15}.

@@

The finer spatial grid for the real population (top left) and commercial data
(top middle) show visually similar patterns, though there are large unsampled
areas in the commercial data from a lack of fishing activity (particularly in
the lower left part of the sampling domain). The survey data at this spatial
resolution shows very sparse information about the spatial distributions of the
populations. The slightly aggregated data on a 5 x 5 grid shows similar
patterns and, while losing some of the spatial detail, there remains good
consistency between the 'real population' and the commercial data. Survey data
starts to pick out some of the similar patterns as the other data sources, but
lacks coverage. The spatial catch information on a 10 x 10 and 20 x 20 grid
loses a significant amount of information about the spatial resolutions for all
data sources, and some differences between the survey, commercial and 'real
population' data emerge. \\

@@

For the commercial data, the most effective closure scenario was based on 1 x 1
data at a monthly temporal resolution. This results in $\sim$ 10 \% reduction
in F for species 1. \textcolor{blue}{This was the only closure scenario to have
	positive effect according to Figure \ref{fig:4}, though looking at the
	trend in Figure \ref{fig:3} this looks more related to the continued
	increased in F trend, as other scenarios had an initial effect}.
Interestingly the monthly data scenario was more effective than weekly data,
which I'd posit is due to the increase amount of data available from the
commercial sampling across a month compared to a week.i Commercial data used at
an annual time-step was ineffective in bringing fishing mortality down for
species 1. \\

Given the scenarios above, it seems clear that spatial disaggregation is more
important than the temporal disaggregation of the commercial data, except when
its used at an annual time-frame, which is the scenario that gave the worst
results.

@@

Our study evaluates the importance of data scaling and considers potential
biases introduced through data aggregation when using commercial fisheries
logbooks and Vessel Monitoring Systems for inference on spatio-temporal
dynamics in fisheries. Understanding how fishers exploit multiple
heterogeneously distributed fish populations with different catch limits or
conservation status requires detailed understanding of the overlap of
resources; this is difficult to achieve using conventional modelling approaches
due to the patchy and irregular nature of fisheries resulting in preferential
sampling \cite{Martinez-Minaya2018}. Often data is aggregated or extrapolated
which requires assumptions about the spatial and temporal scale of processes.
Our study explores the assumptions behind such aggregation and preferential
sampling to identify potential impacts on management advice. With modern
management approaches increasingly employing more nuanced spatio-temporal
approaches in order to maximise productivity while taking account of both the
biological and human processes operating on different time-frames
\cite{Dunn2016}, understanding assumptions behind the data used is vital to
ensure measures are effective. \\

We employ a simulation approach to model each of the population and fishery
processes in a hypothetical 'mixed fishery', allowing us to i) evaluate the
consequences of different aggregation assumptions on our understanding of the
spatio-temporal distribution of the underlying fish populations, and ii)
evaluate the effectiveness of a spatial closure given those assumptions. Our
approach captures fine scale population and fishery dynamics not usually
considered (although see \cite{Bastardie2010, Bailey2018}) which offers the
advantage that larger scale fishery patterns are emergent properties of the
system rather than the result of a statistical modelling framework. \\

Our results show commercial data can provide at right scale and resolution -
depends on scale of process: pop movement etc... Important to consider how
fishers interact / adapt to changes with the resource and mgmt. \\

@@

@@

%%%% figures
\begin{figure}[!ht]
	\includegraphics[width=\linewidth]{../analysis/Data_Aggregation_space_Rev}
	\caption{Data aggregation at different spatial resolutions over a
		ten year period}
	\label{fig:1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width =\linewidth]{../analysis/Data_Aggregation_time_Rev}
	\caption{Data aggregation at different temporal resolutions over a
		three year period}
	\label{fig:2}
\end{figure}
